---
title: "Simulation"
output:   
  pdf_document: default
#  html_document: default
---

# Introduction

our idea:  
women sorted to firms with lower wage premium even in pre period  
women are more likely to receive low premium even within the same firm compared to men  
Covid imposed same firm specific shock on non-wage factorsfor men and women  
(as Covid makes WFH more available, bringing in more flexibility)  
but female further sort out to firms with lower wage premium  
because they value non wage factors than men ? or even more than themselves before Covid?  
(as women are more likely to be the main caregivers at home & caregiving is more critical in the post-Covid period)  

our assumptions:  
1. Men and women operate in different labor markets:  
 different wage premium (normaly lower for women) and value of non-wage factors  
2. Covid shock is likely to have impacted these two markets at the same time but to different extents:  
2.1. worker fixed effect pre/post unchanged: alpha_i  
2.2. firm wage premium pre/post changed: psi  
2.3. firm non-wage factor pre/post changed: amenity  
2.4. female post-COVID non-wage shock is stronger: rho  
2.5. female further sort out to firms with lower wage premium  

# 0. Start
```{r setup, results='hide'}
library(dplyr)
library(tidyr)
library(ggplot2)
library(Matrix)
library(igraph)
library(modelsummary)

set.seed(123)

# Save simulated data to the project folder to avoid re-running every time
project_output_dir <- "D:/master/M2/Applied Labor Economics/final/our code/simulation_rmd"
script_path <- tryCatch(normalizePath(sys.frame(1)$ofile), error = function(e) NA_character_)
output_dir <- if (dir.exists(project_output_dir)) {
    project_output_dir
} else if (!is.na(script_path)) {
    dirname(script_path)
} else {
    getwd()
}
simulation_cache <- file.path(output_dir, "simulated_data_cache.rds")
simulation_cache_version <- 2L
force_resimulate <- FALSE

cache_is_current <- FALSE
if (!force_resimulate && file.exists(simulation_cache)) {
    cached <- readRDS(simulation_cache)
    cache_is_current <- is.list(cached) &&
        isTRUE(cached$version == simulation_cache_version) &&
        all(c("df", "df_simulated") %in% names(cached))
}

if (cache_is_current) {
    df <- cached$df
    df_simulated <- cached$df_simulated
    df <- df %>%
        mutate(period = factor(period, levels = c("pre", "post")))
    df_simulated <- df_simulated %>%
        mutate(period = factor(period, levels = c("pre", "post")))
    print(paste0("Loaded cached simulated data from: ", simulation_cache))
    print(paste0("Cached worker-period rows: ", nrow(df)))
    print(paste0("Cached firm-period rows: ", nrow(df_simulated)))
}
```

# 1. Basic Parameters
```{r parameters, echo=TRUE}
years <- c(2018, 2019, 2021, 2022)
pre_years <- c(2018, 2019)
post_years <- c(2021, 2022)
year_pairs <- data.frame(
    year1 = c(2018, 2021),
    year2 = c(2019, 2022),
    period = c("pre", "post")
)

n_firms <- 1500
n_workers <- 300000
worker_id <- 1:n_workers

offer_shock_sd <- 0.15
full_rank_damping <- 0.85
lambda_grid <- seq(0.02, 0.50, by = 0.01)
sorkin_max_iter <- 10
sorkin_tol <- 1e-4

if (!cache_is_current) {
    female <- rbinom(n_workers, 1, 0.5)
    birth_year <- sample(1965:1995, n_workers, replace = TRUE)

    # worker fixed effects (same for each year)
    alpha_i <- rnorm(n_workers, mean = 0, sd = 0.4)

    # Firm wage premium: (should be estimated with real data from AKM)
    # 1. common firm wage premium base (sd = 0.3)
    psi_base <- scale(rnorm(n_firms))[, 1] * 0.3
    # 2. COVID shock on wage premium (sd = 0.08)
    psi_covid_shock <- scale(rnorm(n_firms))[, 1] * 0.08
    # 3. women more likely to receive lower wage premium within same firm
    female_penalty <- rnorm(n_firms, mean = -0.12, sd = 0.04)

    # Non-wage factors:
    # 1. firm non-wage amenity base: higher = better non-wage value
    amenity_base <- scale(rnorm(n_firms))[, 1]
    # 2. Covid shock on non-wage factors (mostly improve as WFH is more common)
    amenity_covid_shock <- rnorm(n_firms, mean = 0.25, sd = 0.25)
    
    print(paste0("mean of female_penalty:", mean(female_penalty)))
    print(paste0("proportion of women receiving higher wage premium:",
                 mean(female_penalty > 0)))
    print(paste0("mean of amenity_covid_shock:",
                 mean(amenity_covid_shock)))
    print(paste0("proportion of improved amenity:",
                 mean(amenity_covid_shock > 0)))
}
```

# 2. Simulation

```{r Simulation, echo=TRUE}
if (cache_is_current) {
    # The simulation is large.
    # Reuse it unless you do not have simulated_data_cache.rds or change the file name above.
} else {
    print(paste0("Cache path checked: ", simulation_cache))
    print("No current cached simulated data found. Running the simulation now.")

    all_data <- list()
    all_simulated <- list()

    for (p in seq_len(nrow(year_pairs))) {
        year1 <- year_pairs$year1[p]
        year2 <- year_pairs$year2[p]
        period <- year_pairs$period[p]

        post <- ifelse(period == "post", 1, 0)

        age1 <- year1 - birth_year
        age2 <- year2 - birth_year

        exp1 <- age1 - 22 # perfect linearity
        exp2 <- age2 - 22
        
        #--------------------------
        # 2.1. OUR DGP!
        #--------------------------

        # correlation between wage premium and value: after Covid female value non-wage even more
        rho_female <- ifelse(post == 0, 0.6, 0.3)
        rho_male <- ifelse(post == 0, 0.8, 0.7)
        rho <- ifelse(female == 1, rho_female, rho_male)

        # firm-level wage premium
        psi_common <- psi_base + post * psi_covid_shock
        psi_common <- scale(psi_common)[, 1] * 0.3

        # women receive lower premium within same firm
        psi_male <- psi_common
        psi_female <- psi_common + female_penalty

        # common firm-specific non-wage shock for men and women
        amenity <- amenity_base + post * amenity_covid_shock

        # gender-specific latent firm value
        V_male_true <- rho_male * scale(psi_male)[, 1] +
            sqrt(1 - rho_male^2) * scale(amenity)[, 1]

        V_female_true <- rho_female * scale(psi_female)[, 1] +
            sqrt(1 - rho_female^2) * scale(amenity)[, 1]

        #--------------------------
        # 2.2. firm-level objects
        #--------------------------

        lambda1 <- 0.20 # probability an employed worker receives an outside offer (find_lambda.m from Sorkin)
        lambda_ene_base <- 0.08 # probability of an indirect employer change through nonemployment
        sigma_eps <- 1 # scale of idiosyncratic utility shock

        # firm size in the initial assignment: higher value = higher attractiveness = higher firm size
        V_avg_true <- 0.5 * V_male_true + 0.5 * V_female_true
        firm_size <- exp(0.3 * V_avg_true)
        firm_size_prob <- firm_size / sum(firm_size)

        # offer intensity f_j:
        # Large random offer shocks make recovered ranks reflect recruiting intensity,
        # not only worker-valued firm quality.
        offer_shock <- rnorm(n_firms, 0, offer_shock_sd)
        offer <- exp(0.5 * V_avg_true + offer_shock)
        offer_prob <- offer / sum(offer)

        #--------------------------
        # 2.3. initial firms
        #--------------------------

        current_firm <- sample(1:n_firms, size = n_workers, replace = TRUE, prob = firm_size_prob)

        psi_current <- ifelse(
            female == 1,
            psi_female[current_firm],
            psi_male[current_firm]
        )

        # baseline wage in year 1: 10 + worker fixed effect + firm fixed effect + age and its polinomials (without exp1) + non-wage factor
        logy1 <- 10 + alpha_i + psi_current +
            0.03 * age1 - 0.0002 * age1^2 + 0.00001 * age1^3 +
            rnorm(n_workers, 0, 0.2)

        #--------------------------
        # 2.4. outside offers and job-to-job decisions
        #--------------------------

        gets_offer <- rbinom(n_workers, size = 1, prob = lambda1)

        # offered firm
        offer_firm <- sample(1:n_firms, size = n_workers, replace = TRUE, prob = offer_prob)

        # prevent same-firm "offers" by redrawing when necessary
        same_firm <- which(offer_firm == current_firm)
        while (length(same_firm) > 0) {
            offer_firm[same_firm] <- sample(1:n_firms, size = length(same_firm), replace = TRUE, prob = offer_prob)
            same_firm <- which(offer_firm == current_firm)
        }
        # EE -- Direct employer-to-employer move:
        # workers compare utility current vs utility offer: U = common firm value + iid shock
        V_current <- ifelse(
            female == 1,
            V_female_true[current_firm],
            V_male_true[current_firm]
        )

        V_offer <- ifelse(
            female == 1,
            V_female_true[offer_firm],
            V_male_true[offer_firm]
        )

        eps_current <- rnorm(n_workers, 0, sigma_eps)
        eps_offer <- rnorm(n_workers, 0, sigma_eps)

        U_current <- V_current + eps_current
        U_offer <- V_offer + eps_offer

        # revealed preferences
        move <- ifelse(gets_offer == 1 & U_offer > U_current, 1, 0)

        # ENE -- Indirect employer change through nonemployment:

        # Lower-valued firms have higher separation risk
        ene_prob <- plogis(qlogis(lambda_ene_base) - 0.35 * V_current)
        ene <- as.integer(move == 0 & rbinom(n_workers, size = 1, prob = ene_prob) == 1)

        # separated workers draw their next employer from the same offer-arrival distribution.
        ene_firm <- sample(1:n_firms, size = n_workers, replace = TRUE, prob = offer_prob)
        same_firm_ene <- which(ene == 1 & ene_firm == current_firm)
        while (length(same_firm_ene) > 0) {
            ene_firm[same_firm_ene] <- sample(
                1:n_firms,
                size = length(same_firm_ene),
                replace = TRUE,
                prob = offer_prob
            )
            same_firm_ene <- which(ene == 1 & ene_firm == current_firm)
        }

        next_firm <- ifelse(move == 1, offer_firm, ifelse(ene == 1, ene_firm, current_firm))

        psi_next <- ifelse(
            female == 1,
            psi_female[next_firm],
            psi_male[next_firm]
        )

        # year 2 wages
        logy2 <- 10 + alpha_i + psi_next +
            0.03 * age2 - 0.0002 * age2^2 + 0.00001 * age2^3 +
            rnorm(n_workers, 0, 0.2)

        # EE indicator: 1 if changed employer directly
        ee <- as.integer(move == 1)

        # ENE indicator: 1 if changed employer indirectly via nonemployment
        ene <- as.integer(ene == 1)

        #--------------------------
        # 2.5. simulated dataset
        #--------------------------

        matrix_sim <- data.frame(
            worker_id = worker_id,
            female = female,
            year1 = year1,
            year2 = year2,
            period = period,
            age1 = age1,
            age2 = age2,
            index1 = current_firm,
            index2 = next_firm,
            logy1 = logy1,
            logy2 = logy2,
            ee = ee,
            ene = ene
        )

        simulated <- data.frame(
            firm_id = 1:n_firms,
            period = period,
            psi_male = psi_male,
            psi_female = psi_female,
            amenity = amenity,
            female_penalty = female_penalty,
            rho_male = rho_male,
            rho_female = rho_female,
            V_male_true = V_male_true,
            V_female_true = V_female_true
        )

        all_data[[p]] <- matrix_sim
        all_simulated[[p]] <- simulated
    }

    df <- bind_rows(all_data) #
    df_simulated <- bind_rows(all_simulated)

    # Keep the period ordering chronological in all tables and figures.
    # Otherwise ggplot orders character labels alphabetically, putting "post"
    # before "pre".
    df <- df %>%
        mutate(period = factor(period, levels = c("pre", "post")))
    df_simulated <- df_simulated %>%
        mutate(period = factor(period, levels = c("pre", "post")))

    saveRDS(
        list(
            version = simulation_cache_version,
            df = df,
            df_simulated = df_simulated
        ),
        simulation_cache
    )
    print(paste0("Saved simulated data cache to: ", simulation_cache))
}

stopifnot(all(df$ee + df$ene <= 1))

sorkin_empty_rank <- function(period_keep, gender_keep, method_keep) {
    data.frame(
        firm_id = integer(),
        period = character(),
        female = integer(),
        method = character(),
        V_hat = numeric()
    )
}

# Stops early if an estimator produces too few usable observations
require_rows <- function(data, min_rows, label) {
    if (nrow(data) < min_rows) {
        stop(paste0(
            label,
            " has only ",
            nrow(data),
            " usable rows. Check mobility rates or estimator connectivity before plotting."
        ))
    }
    invisible(data)
}

# Builds the directed mobility matrix used by PageRank
# Columns are origin firms and rows are destination firms
# matches the Sorkin code: M[j, k] is the number of moves from k to j.
build_count_matrix <- function(origin, dest, n_nodes) {
    if (length(origin) == 0) {
        return(matrix(0, nrow = n_nodes, ncol = n_nodes))
    }

    # M[dest, origin] counts moves from origin to destination.
    # sparseMatrix automatically sums duplicate origin-destination pairs.
    as.matrix(sparseMatrix(
        i = dest,
        j = origin,
        x = 1,
        dims = c(n_nodes, n_nodes)
    ))
}

# Standardizes a vector when it has variation; otherwise only recenters it
# This keeps constants from becoming NA when sd(x) is zero.
standardize_or_center <- function(x) {
    sx <- sd(x, na.rm = TRUE)
    if (is.na(sx) || sx == 0) {
        return(as.numeric(x - mean(x, na.rm = TRUE)))
    }
    as.numeric((x - mean(x, na.rm = TRUE)) / sx)
}

# Converts a noisy cardinal score into a normal-score ranking.
# This is useful for PageRank/eigenvector-style firm values:
# the rank order is the identifying information,
# while raw eigenvector magnitudes can explode in finite simulated mobility graphs with weakly connected firms.
rank_to_normal_score <- function(x) {
    ok <- is.finite(x) & !is.na(x)
    out <- rep(NA_real_, length(x))
    n_ok <- sum(ok)
    if (n_ok < 2) {
        out[ok] <- 0
        return(out)
    }
    r <- rank(x[ok], ties.method = "average")
    p <- (r - 0.5) / n_ok
    out[ok] <- qnorm(p)
    standardize_or_center(out)
}

# Small helper: escapes underscores for laTeX
escape_latex <- function(x) {
    x <- as.character(x)
    gsub("_", "\\\\_", x, fixed = TRUE)
}

# LaTeX table
write_simple_latex_table <- function(tab, path, caption = NULL, digits = 3) {
    tab_out <- tab
    for (nm in names(tab_out)) {
        if (is.numeric(tab_out[[nm]])) {
            tab_out[[nm]] <- ifelse(
                is.na(tab_out[[nm]]),
                "",
                formatC(tab_out[[nm]], digits = digits, format = "f")
            )
        } else {
            tab_out[[nm]] <- escape_latex(tab_out[[nm]])
        }
    }
    lines <- c(
        "\\begin{table}[!htbp]\\centering",
        if (!is.null(caption)) paste0("\\caption{", escape_latex(caption), "}") else NULL,
        paste0("\\begin{tabular}{", paste(rep("l", ncol(tab_out)), collapse = ""), "}"),
        "\\hline",
        paste(escape_latex(names(tab_out)), collapse = " & "),
        "\\\\ \\hline"
    )
    body <- apply(tab_out, 1, function(z) paste(z, collapse = " & "))
    lines <- c(lines, paste0(body, " \\\\"), "\\hline", "\\end{tabular}", "\\end{table}")
    writeLines(lines, path)
}

# Finds the largest connected part of the mobility graph.
# The strict Sorkin estimator uses the largest strongly connected component;
# weak connectivity is only used as a finite-simulation fallback.
largest_component_nodes <- function(M, strong = TRUE) {
    g <- graph_from_adjacency_matrix(
        t(M),
        mode = "directed",
        weighted = TRUE,
        diag = FALSE
    )

    scc <- components(g, mode = ifelse(strong, "strong", "weak"))
    which(scc$membership == which.max(scc$csize))
}

# EE: job-to-job moves only
# extracts the eigenvector from the largest connected mobility graph.
recover_sorkin_value <- function(M, node_ids, min_nodes = 10) {
    diag(M) <- 0

    # Sorkin choice: largest strongly connected component.
    # (In small simulated subsamples this can occasionally be too small,
    # especially when many transitions pass through nonemployment. )
    # Our fallback: largest weakly connected active component & tiny pseudo-flow

    keep <- largest_component_nodes(M, strong = TRUE)
    active <- which(rowSums(M) > 0 & colSums(M) > 0)
    used_weak_fallback <- FALSE

    if (length(intersect(keep, active)) < min_nodes) {
        keep <- intersect(largest_component_nodes(M, strong = FALSE), active)
        used_weak_fallback <- TRUE
        message(
            "Using weak-component fallback because the strict strongly connected ",
            "set had fewer than ", min_nodes, " active nodes."
        )
    }

    if (length(keep) < 2) {
        return(data.frame(node_id = integer(), V_hat = numeric()))
    }

    M_sub <- M[keep, keep, drop = FALSE]
    node_ids_sub <- node_ids[keep]

    if (used_weak_fallback) {
        pseudo_flow <- 1e-6
        M_sub <- M_sub + pseudo_flow
        diag(M_sub) <- 0
    }

    # M_sub[dest, origin] stores moves from origin to destination (tilde_exptV.m)
    # The normalization follows tilde_exptV.m: S = diag(sum(M_sub));
    # A = solve(S) %*% M_sub.
    outflow <- colSums(M_sub)
    keep_positive <- which(outflow > 0)

    if (length(keep_positive) < 2) {
        return(data.frame(node_id = integer(), V_hat = numeric()))
    }

    M_sub <- M_sub[keep_positive, keep_positive, drop = FALSE]
    node_ids_sub <- node_ids_sub[keep_positive]
    S_sub <- diag(colSums(M_sub))
    A <- solve(S_sub) %*% M_sub

    eig <- eigen(A)
    idx <- which.max(Re(eig$values))

    q <- abs(Re(eig$vectors[, idx]))
    q <- pmax(q, 1e-10)

    # The eigenvector is positive only up to scale.
    # We use a rank-normalized version
    # so one tiny eigenvector cell cannot dominate the graph.
    data.frame(
        node_id = node_ids_sub,
        V_hat = rank_to_normal_score(log(q))
    )
}

# EE+ENE
# the ranking remains defined when some firms have sparse flows
recover_augmented_value <- function(M, node_ids, damping = full_rank_damping) {
    diag(M) <- 0
    active <- which(rowSums(M) > 0 | colSums(M) > 0)

    if (length(active) < 2) {
        return(data.frame(node_id = integer(), V_hat = numeric()))
    }

    M_sub <- M[active, active, drop = FALSE]
    node_ids_sub <- node_ids[active]
    outflow <- colSums(M_sub)
    n <- nrow(M_sub)

    # Full EE+ENE matrices can be weakly connected because nonemployment is a hub.
    # To keep the estimator usable in simulation, we use a damped transition matrix:
    # columns with no outflow send equal mass everywhere, and every column receives
    # a small teleportation probability. This is a stable approximation to the
    # full mobility ranking, while the EE-only option above keeps the stricter
    # Sorkin eigenvector.
    A <- matrix(1 / n, nrow = n, ncol = n)
    positive_outflow <- which(outflow > 0)
    A[, positive_outflow] <- sweep(
        M_sub[, positive_outflow, drop = FALSE],
        2,
        outflow[positive_outflow],
        "/"
    )

    A <- damping * A + (1 - damping) * matrix(1 / n, nrow = n, ncol = n)

    eig <- eigen(A)
    idx <- which.max(Re(eig$values))
    q <- abs(Re(eig$vectors[, idx]))
    q <- pmax(q / sum(q), 1e-10)

    # Damped PageRank score from the full EE+ENE graph. Rank-normalization keeps
    # the plot and regressions about ordering, not arbitrary eigenvector scale.
    data.frame(
        node_id = node_ids_sub,
        V_hat = rank_to_normal_score(log(q))
    )
}

# Stable validation ranking based on net inflows. It is less structural than
# Sorkin's full model, but very useful for diagnosing simulated mobility data.
recover_full_flow_score <- function(data, period_keep, gender_keep) {
    d <- data %>%
        filter(period == period_keep)

    if (!is.null(gender_keep)) {
        d <- d %>% filter(female == gender_keep)
    }

    # Full-flow validation score:
    # - EE and ENE destinations are evidence that a firm is attractive.
    # - EE and ENE origins are evidence that workers leave that firm.
    # - The +1 smoothing keeps every firm in the graph and prevents log(0).
    # This is intentionally simpler and more stable than the structural Sorkin
    # loop. It is the right object for checking whether the simulated mobility
    # patterns point toward the true latent firm value.
    transition_rows <- d %>%
        filter((ee == 1 | ene == 1), index1 != index2)

    inflow <- transition_rows %>%
        count(index2, name = "inflow") %>%
        rename(firm_id = index2)

    outflow <- transition_rows %>%
        count(index1, name = "outflow") %>%
        rename(firm_id = index1)

    data.frame(firm_id = 1:n_firms) %>%
        left_join(inflow, by = "firm_id") %>%
        left_join(outflow, by = "firm_id") %>%
        mutate(
            inflow = replace_na(inflow, 0),
            outflow = replace_na(outflow, 0),
            V_hat = rank_to_normal_score(log((inflow + 1) / (outflow + 1))),
            period = period_keep,
            female = gender_keep,
            method = "full"
        ) %>%
        select(firm_id, period, female, method, V_hat)
}

# Constructs the augmented firm/nonemployment transition matrix and the objects
# needed to recover lambda and employer value in the Sorkin-style model.
build_sorkin_inputs <- function(data, period_keep, gender_keep) {
    d <- data %>%
        filter(period == period_keep)

    if (!is.null(gender_keep)) {
        d <- d %>% filter(female == gender_keep)
    }

    ee_rows <- d %>% filter(ee == 1, index1 != index2)
    ene_rows <- d %>% filter(ene == 1, index1 != index2)

    nonemployment_node <- 1L
    n_nodes <- n_firms + 1L

    origin <- c(
        ee_rows$index1 + 1L,
        ene_rows$index1 + 1L,
        rep(nonemployment_node, nrow(ene_rows))
    )
    dest <- c(
        ee_rows$index2 + 1L,
        rep(nonemployment_node, nrow(ene_rows)),
        ene_rows$index2 + 1L
    )

    M <- build_count_matrix(origin = origin, dest = dest, n_nodes = n_nodes)

    # g_level is firm employment/person-years. Node 1 is nonemployment.
    g_level <- numeric(n_nodes)
    origin_counts <- tabulate(d$index1 + 1L, nbins = n_nodes)
    g_level <- g_level + origin_counts

    # fo_level is hires from nonemployment, i.e. ENE destinations.
    fo_level <- numeric(n_nodes)
    fo_level <- fo_level + tabulate(ene_rows$index2 + 1L, nbins = n_nodes)
    fo_level[nonemployment_node] <- max(1, sum(fo_level))

    list(
        data = d,
        ee_rows = ee_rows,
        ene_rows = ene_rows,
        M = M,
        g_level = g_level,
        fo_level = fo_level,
        ee_count = nrow(ee_rows),
        ene_count = nrow(ene_rows),
        nonemployment_node = nonemployment_node
    )
}

# Recovers exp(V) from the augmented transition matrix using a damped
# eigenvector/PageRank step.
recover_exptv_damped <- function(M, damping = full_rank_damping) {
    diag(M) <- 0
    n <- nrow(M)
    outflow <- colSums(M)

    # Sorkin's tilde_exptV solves the eigenvector of a column-normalized mobility
    # matrix. The damping term keeps simulated finite samples from collapsing when
    # the mobility graph is not fully strongly connected.
    A <- matrix(1 / n, nrow = n, ncol = n)
    positive_outflow <- which(outflow > 0)
    A[, positive_outflow] <- sweep(
        M[, positive_outflow, drop = FALSE],
        2,
        outflow[positive_outflow],
        "/"
    )
    A <- damping * A + (1 - damping) * matrix(1 / n, nrow = n, ncol = n)

    eig <- eigen(A)
    idx <- which.max(Re(eig$values))
    exptV <- abs(Re(eig$vectors[, idx]))
    pmax(exptV / sum(exptV), 1e-10)
}

# Given a candidate lambda, recovers employer value and the implied offer
# distribution, then compares the model-implied EE rate with the data.
sorkin_given_lambda <- function(M, exptV, g_level, fo_level, lambda1, n_groups = 200) {
    eps <- 1e-8
    g_level <- pmax(g_level, 0)
    fo_level <- pmax(fo_level, 0)

    # Smooth the nonemployment-hire distribution so b = exp(V) * g / fo is finite
    # for every firm. This is a finite-simulation safeguard; with large real data,
    # Sorkin restricts to firms with nonemployment hires.
    fo <- (fo_level + eps) / sum(fo_level + eps)
    g <- (g_level + eps) / sum(g_level + eps)
    g_ns <- g

    hires_ne <- sum(fo_level)
    W <- sum(g_level)
    if (hires_ne <= 0 || W <= 0) {
        return(NULL)
    }

    b <- (exptV * g_ns) / fo
    a <- (exptV[1] * hires_ne) / W
    a2 <- a / (1 - lambda1)
    CexpV <- b - a2

    positive <- CexpV > 0 & is.finite(CexpV)
    if (!any(positive)) {
        return(NULL)
    }
    CexpV[!positive] <- min(CexpV[positive])

    fC <- fo * (b / CexpV)
    C1 <- 1 / sum(fC[-1])
    f <- fC[-1] * C1

    expVe <- CexpV[-1] / C1
    expVn_init <- a2 / C1
    expVe <- expVe / expVn_init
    Ve <- log(pmax(expVe, eps))
    Ve_raw <- Ve

    firm_g <- g[-1]
    firm_g <- firm_g / sum(firm_g)
    f <- pmax(f, eps)
    f <- f / sum(f)

    # Match lambda by comparing the model-implied EE rate to the observed EE rate.
    # Sorkin groups firms by V_e for speed. We do the same, with fewer groups for R.
    ord <- order(Ve)
    Ve2 <- Ve[ord]
    g2 <- firm_g[ord]
    f2 <- f[ord]
    cum_g <- cumsum(g2)
    group <- pmin(n_groups, floor(cum_g * n_groups) + 1L)

    grouped <- data.frame(group = group, Ve = Ve2, g = g2, f = f2) %>%
        group_by(group) %>%
        summarise(
            Ve = weighted.mean(Ve, g),
            g = sum(g),
            f = sum(f),
            .groups = "drop"
        )

    P <- outer(
        exp(grouped$Ve),
        exp(grouped$Ve),
        FUN = function(i, j) i / (i + j)
    )
    diag(P) <- 0
    prob_accept_model <- mean(as.numeric(t(grouped$f) %*% P))
    ee_prob_model <- lambda1 * prob_accept_model

    ee_prob_data <- sum(M[-1, -1]) / sum(g_level[-1])
    gap <- abs(ee_prob_model - ee_prob_data) / max(ee_prob_data, eps)

    list(
        gap = gap,
        Ve = rank_to_normal_score(Ve_raw),
        Ve_raw = Ve_raw,
        f = f,
        C1 = C1,
        ee_prob_model = ee_prob_model,
        ee_prob_data = ee_prob_data
    )
}

# Searches over lambda values and keeps the one whose model-implied EE rate is
# closest to the observed EE rate.
estimate_lambda_and_ve <- function(M, g_level, fo_level, grid = lambda_grid) {
    exptV <- recover_exptv_damped(M)
    fits <- lapply(grid, function(lambda1) {
        fit <- sorkin_given_lambda(M, exptV, g_level, fo_level, lambda1)
        if (is.null(fit)) {
            return(data.frame(lambda = lambda1, gap = Inf))
        }
        data.frame(
            lambda = lambda1,
            gap = fit$gap,
            ee_prob_model = fit$ee_prob_model,
            ee_prob_data = fit$ee_prob_data
        )
    }) %>%
        bind_rows()

    best_lambda <- fits$lambda[which.min(fits$gap)]
    best_fit <- sorkin_given_lambda(M, exptV, g_level, fo_level, best_lambda)

    list(
        lambda = best_lambda,
        grid = fits,
        Ve = best_fit$Ve,
        f = best_fit$f,
        C1 = best_fit$C1,
        exptV = exptV,
        ee_prob_data = best_fit$ee_prob_data,
        ee_prob_model = best_fit$ee_prob_model
    )
}

# Converts recovered value, offer probabilities, and lambda into model-consistent
# EE and EN transition rates.
model_consistent_rates <- function(Ve, f, lambda1, g_level) {
    eps <- 1e-8
    n_firms_local <- length(Ve)
    expVe <- exp(Ve)
    f <- pmax(f, eps)
    f <- f / sum(f)

    # Probability a worker at firm j accepts an EE offer from the offer distribution.
    # P[i, j] is acceptance probability for destination i from origin j.
    P <- outer(expVe, expVe, FUN = function(i, j) i / (i + j))
    diag(P) <- 0
    prob_accept_ee <- as.numeric(t(f) %*% P)
    ee_counterfactual <- lambda1 * pmin(pmax(prob_accept_ee, 0), 1)

    # Probability of EN separation/nonemployment acceptance. Vn is normalized to 0.
    en_counterfactual <- (1 - lambda1) * (1 / (expVe + 1))
    en_counterfactual <- pmin(pmax(en_counterfactual, 0), 1)

    # Avoid exactly zero weights, which can disconnect the finite simulated graph.
    list(
        ee = pmax(ee_counterfactual, eps),
        en = pmax(en_counterfactual, eps)
    )
}

# Reweights the observed transition graph using the model-consistent EE and EN
# rates from the previous iteration.
compute_weighted_M <- function(inputs, ee_counterfactual, en_counterfactual) {
    n_nodes <- n_firms + 1L
    nonemployment_node <- 1L

    ee_rows <- inputs$ee_rows
    ene_rows <- inputs$ene_rows

    origin <- c(
        ee_rows$index1 + 1L,
        ene_rows$index1 + 1L,
        rep(nonemployment_node, nrow(ene_rows))
    )
    dest <- c(
        ee_rows$index2 + 1L,
        rep(nonemployment_node, nrow(ene_rows)),
        ene_rows$index2 + 1L
    )
    weights <- c(
        ee_counterfactual[ee_rows$index1],
        en_counterfactual[ene_rows$index1],
        rep(1, nrow(ene_rows))
    )

    if (length(origin) == 0) {
        return(matrix(0, nrow = n_nodes, ncol = n_nodes))
    }

    as.matrix(sparseMatrix(
        i = dest,
        j = origin,
        x = weights,
        dims = c(n_nodes, n_nodes)
    ))
}

# Iterates between lambda/value recovery and transition-matrix reweighting until
# the recovered employer values stabilize.
estimate_sorkin_iterative <- function(inputs, grid = lambda_grid,
                                      max_iter = sorkin_max_iter,
                                      tol = sorkin_tol) {
    M_current <- inputs$M
    previous_Ve <- NULL
    history <- list()

    for (iter in seq_len(max_iter)) {
        fit <- estimate_lambda_and_ve(M_current, inputs$g_level, inputs$fo_level, grid)
        rates <- model_consistent_rates(fit$Ve, fit$f, fit$lambda, inputs$g_level)

        if (is.null(previous_Ve)) {
            corr_old_new <- NA_real_
            gap <- Inf
        } else {
            corr_old_new <- suppressWarnings(cor(previous_Ve, fit$Ve, use = "complete.obs"))
            gap <- mean(abs(previous_Ve - fit$Ve), na.rm = TRUE)
        }

        history[[iter]] <- data.frame(
            iteration = iter,
            lambda_hat = fit$lambda,
            ee_prob_data = fit$ee_prob_data,
            ee_prob_model = fit$ee_prob_model,
            mean_abs_change = gap,
            corr_old_new = corr_old_new
        )

        if (!is.null(previous_Ve) && is.finite(gap) && gap < tol) {
            return(list(
                fit = fit,
                history = bind_rows(history),
                M_final = M_current,
                ee_counterfactual = rates$ee,
                en_counterfactual = rates$en,
                converged = TRUE
            ))
        }

        previous_Ve <- fit$Ve
        M_current <- compute_weighted_M(inputs, rates$ee, rates$en)
    }

    list(
        fit = fit,
        history = bind_rows(history),
        M_final = M_current,
        ee_counterfactual = rates$ee,
        en_counterfactual = rates$en,
        converged = FALSE
    )
}

# Full Sorkin-style recovery for one gender-period cell. It saves lambda and
# iteration diagnostics, and falls back to the stable ranking only if the finite
# simulated graph produces a degenerate cardinal value vector.
recover_sorkin_full_model <- function(data, period_keep, gender_keep) {
    inputs <- build_sorkin_inputs(data, period_keep, gender_keep)
    if ((inputs$ee_count + inputs$ene_count) == 0) {
        return(sorkin_empty_rank(period_keep, gender_keep, "sorkin_full"))
    }

    iter_fit <- estimate_sorkin_iterative(inputs)
    fit <- iter_fit$fit
    iteration_history <- iter_fit$history %>%
        mutate(period = period_keep, female = gender_keep)
    print(iteration_history)

    history_path <- file.path(
        output_dir,
        paste0("sorkin_iteration_history_", period_keep, "_female", gender_keep, ".csv")
    )
    write.csv(iteration_history, history_path, row.names = FALSE)

    print(data.frame(
        period = period_keep,
        female = gender_keep,
        lambda_hat = fit$lambda,
        ee_prob_data = fit$ee_prob_data,
        ee_prob_model = fit$ee_prob_model,
        ee_moves = inputs$ee_count,
        ene_moves = inputs$ene_count,
        converged = iter_fit$converged,
        iterations = nrow(iter_fit$history)
    ))

    value_out <- data.frame(
        firm_id = 1:n_firms,
        period = period_keep,
        female = gender_keep,
        method = "sorkin_full",
        lambda_hat = fit$lambda,
        offer_hat = fit$f,
        # fit$Ve is already rank-normalized inside sorkin_given_lambda().
        # Applying the transformation again is harmless and protects the rest
        # of the script if that internal normalization is changed later.
        V_hat = rank_to_normal_score(fit$Ve)
    )

    bad_value <- sum(is.finite(value_out$V_hat)) < 10 ||
        sd(value_out$V_hat, na.rm = TRUE) == 0 ||
        max(abs(value_out$V_hat), na.rm = TRUE) > 4

    if (bad_value) {
        warning(
            "Sorkin full value was degenerate for period=", period_keep,
            ", female=", gender_keep,
            ". Falling back to the stable full-flow PageRank score for V_hat, ",
            "while keeping lambda_hat from the Sorkin step."
        )
        value_out <- recover_full_flow_score(data, period_keep, gender_keep) %>%
            mutate(
                method = "sorkin_full_rank_fallback",
                lambda_hat = fit$lambda,
                offer_hat = fit$f
            )
    }

    value_out
}

```

# 3. revealed-preference ranking function
```{r function, echo=TRUE}
# Public wrapper choosing which revealed-preference ranking to estimate:
# EE-only, stable full-flow, or full Sorkin-style EE+ENE.
estimate_rank <- function(data, period_keep, gender_keep = NULL, method_keep = "ee") {
    d <- data %>%
        filter(period == period_keep)

    if (!is.null(gender_keep)) {
        d <- d %>% filter(female == gender_keep)
    }

    if (!method_keep %in% c("ee", "full", "sorkin_full")) {
        stop("method_keep must be 'ee', 'full', or 'sorkin_full'")
    }

    #--------------------------
    # 3.1 construct mobility matrix M[j, k] = moves from k to j
    #--------------------------

    ee_rows <- d %>% filter(ee == 1, index1 != index2)
    ene_rows <- d %>% filter(ene == 1, index1 != index2)

    print(paste0(
        "Period: ", period_keep,
        ", gender: ", gender_keep,
        ", method: ", method_keep,
        ", Number of EE moves: ", nrow(ee_rows),
        ", Number of ENE moves: ", nrow(ene_rows)
    ))

    if (method_keep == "ee" && nrow(ee_rows) == 0) {
        return(sorkin_empty_rank(period_keep, gender_keep, method_keep))
    }

    if (method_keep == "full" && (nrow(ee_rows) + nrow(ene_rows)) == 0) {
        return(sorkin_empty_rank(period_keep, gender_keep, method_keep))
    }

    if (method_keep == "sorkin_full" && (nrow(ee_rows) + nrow(ene_rows)) == 0) {
        return(sorkin_empty_rank(period_keep, gender_keep, method_keep))
    }

    if (method_keep == "ee") {
        node_ids <- 1:n_firms
        M <- build_count_matrix(
            origin = ee_rows$index1,
            dest = ee_rows$index2,
            n_nodes = n_firms
        )
        recovered <- recover_sorkin_value(M, node_ids)
        return(recovered %>%
            transmute(
                firm_id = node_id,
                period = period_keep,
                female = gender_keep,
                method = method_keep,
                V_hat = V_hat
            ))
    }

    if (method_keep == "sorkin_full") {
        return(recover_sorkin_full_model(data, period_keep, gender_keep))
    }

    recover_full_flow_score(data, period_keep, gender_keep)
}
```

# 5. Estimation
```{r estimation, echo=TRUE}
#--------------------------
# 5.0 data preparation & summary
#--------------------------

wage_long <- bind_rows(
    df %>%
        transmute(
            worker_id,
            female,
            period,
            year = year1,
            age = age1,
            firm_id = index1,
            logy = logy1
        ),
    df %>%
        transmute(
            worker_id,
            female,
            period,
            year = year2,
            age = age2,
            firm_id = index2,
            logy = logy2
        )
) %>%
    mutate(
        female = as.integer(female),
        gender = ifelse(female == 1, "Female", "Male"),
        post = as.integer(period == "post")
    )

# Estimate AKM-style firm effects separately within each gender-period cell.
# The model is log wage = worker FE + firm FE + year FE + age controls + error.
# To avoid requiring large external fixed-effect packages, the code first removes
# age and year effects, then alternates between worker and firm means until the
# two sets of fixed effects stabilize. The firm effects are centered to have
# employment-weighted mean zero inside each gender-period cell.
estimate_akm_firm_premia <- function(data_subset, max_iter = 200, tol = 1e-7) {
    require_rows(data_subset, min_rows = 10, label = "AKM subset")

    base_resid <- resid(lm(logy ~ age + I(age^2) + I(age^3) + factor(year), data = data_subset))
    worker_levels <- sort(unique(data_subset$worker_id))
    firm_levels <- sort(unique(data_subset$firm_id))

    worker_index <- match(data_subset$worker_id, worker_levels)
    firm_index <- match(data_subset$firm_id, firm_levels)

    alpha <- numeric(length(worker_levels))
    psi <- numeric(length(firm_levels))

    for (iter in seq_len(max_iter)) {
        old_psi <- psi

        alpha <- rowsum(base_resid - psi[firm_index], worker_index, reorder = FALSE)[, 1] /
            as.vector(tabulate(worker_index, nbins = length(worker_levels)))

        psi <- rowsum(base_resid - alpha[worker_index], firm_index, reorder = FALSE)[, 1] /
            as.vector(tabulate(firm_index, nbins = length(firm_levels)))

        firm_weights <- as.vector(tabulate(firm_index, nbins = length(firm_levels)))
        psi <- psi - weighted.mean(psi, firm_weights)

        change <- max(abs(psi - old_psi), na.rm = TRUE)
        if (is.finite(change) && change < tol) {
            break
        }
    }

    data.frame(
        firm_id = firm_levels,
        psi_hat = as.numeric(psi),
        n_obs = as.vector(tabulate(firm_index, nbins = length(firm_levels)))
    )
}

# Run the AKM estimator four times, matching the report: male-pre, male-post,
# female-pre, and female-post are treated as separate labor markets.
akm_firm_premia_long <- wage_long %>%
    group_by(period, female) %>%
    group_modify(~ estimate_akm_firm_premia(.x)) %>%
    ungroup() %>%
    mutate(
        gender = ifelse(female == 1, "Female", "Male")
    )

write.csv(
    akm_firm_premia_long,
    file.path(output_dir, "akm_firm_premia_gender_period.csv"),
    row.names = FALSE
)

summary_table <- wage_long %>%
    group_by(gender, period) %>%
    summarise(
        n_obs = n(),
        n_workers = n_distinct(worker_id),
        n_firms = n_distinct(firm_id),
        mean_log_wage = mean(logy, na.rm = TRUE),
        sd_log_wage = sd(logy, na.rm = TRUE),
        p10_log_wage = quantile(logy, 0.10, na.rm = TRUE),
        p50_log_wage = quantile(logy, 0.50, na.rm = TRUE),
        p90_log_wage = quantile(logy, 0.90, na.rm = TRUE),
        mean_age = mean(age, na.rm = TRUE),
        sd_age = sd(age, na.rm = TRUE),
        .groups = "drop"
    )

# One consolidated descriptive table by year and gender.
# EE/ENE moves are attached to the origin year of the transition: in the current
# annual simulation, this means 2018 for pre moves and 2021 for post moves.
true_firm_premia_long <- bind_rows(
    df_simulated %>%
        transmute(
            firm_id,
            period,
            gender = "Male",
            psi_true = psi_male
        ),
    df_simulated %>%
        transmute(
            firm_id,
            period,
            gender = "Female",
            psi_true = psi_female
        )
)

mobility_by_year_gender <- df %>%
    mutate(
        year = year1,
        gender = ifelse(female == 1, "Female", "Male")
    ) %>%
    group_by(year, gender) %>%
    summarise(
        num_ee = sum(ee, na.rm = TRUE),
        num_ene = sum(ene, na.rm = TRUE),
        ee_rate = mean(ee, na.rm = TRUE),
        ene_rate = mean(ene, na.rm = TRUE),
        .groups = "drop"
    )

data_summary_by_year_gender <- wage_long %>%
    left_join(
        akm_firm_premia_long %>% select(period, female, firm_id, psi_hat),
        by = c("period", "female", "firm_id")
    ) %>%
    left_join(
        true_firm_premia_long,
        by = c("period", "gender", "firm_id")
    ) %>%
    group_by(year, gender) %>%
    summarise(
        num_obs = n(),
        num_workers = n_distinct(worker_id),
        num_firms = n_distinct(firm_id),
        mean_log_wage = mean(logy, na.rm = TRUE),
        sd_log_wage = sd(logy, na.rm = TRUE),
        mean_akm_psi = mean(psi_hat, na.rm = TRUE),
        sd_akm_psi = sd(psi_hat, na.rm = TRUE),
        p10_akm_psi = quantile(psi_hat, 0.10, na.rm = TRUE),
        p50_akm_psi = quantile(psi_hat, 0.50, na.rm = TRUE),
        p90_akm_psi = quantile(psi_hat, 0.90, na.rm = TRUE),
        mean_true_psi = mean(psi_true, na.rm = TRUE),
        sd_true_psi = sd(psi_true, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    left_join(mobility_by_year_gender, by = c("year", "gender")) %>%
    mutate(
        num_ee = replace_na(num_ee, 0),
        num_ene = replace_na(num_ene, 0),
        ee_rate = replace_na(ee_rate, 0),
        ene_rate = replace_na(ene_rate, 0)
    ) %>%
    arrange(year, gender)

print(data_summary_by_year_gender)

write.csv(
    data_summary_by_year_gender,
    file.path(output_dir, "data_summary_by_year_gender.csv"),
    row.names = FALSE
)

write_simple_latex_table(
    data_summary_by_year_gender,
    file.path(output_dir, "data_summary_by_year_gender_full.tex"),
    caption = "Summary Statistics by Year and Gender"
)

data_summary_by_year_gender_neat <- data_summary_by_year_gender %>%
    transmute(
        year = as.character(as.integer(year)),
        gender,
        workers = as.character(as.integer(num_workers)),
        firms = as.character(as.integer(num_firms)),
        EE = as.character(as.integer(num_ee)),
        ENE = as.character(as.integer(num_ene)),
        `EE rate` = sprintf("%.3f", ee_rate),
        `ENE rate` = sprintf("%.3f", ene_rate),
        `Mean log wage` = sprintf("%.3f", mean_log_wage),
        `Mean AKM psi` = sprintf("%.3f", mean_akm_psi),
        `Mean true psi` = sprintf("%.3f", mean_true_psi)
    )

write_simple_latex_table(
    data_summary_by_year_gender_neat,
    file.path(output_dir, "data_summary_by_year_gender.tex"),
    caption = "Summary Statistics by Year and Gender"
)


modelsummary::datasummary_df(
    summary_table,
    output = file.path(output_dir, "summary_statistics.tex"),
    title = "Summary Statistics by Gender and Period"
)

firm_gap <- df_simulated %>%
    mutate(
        post = as.integer(period == "post"),
        psi_gap = psi_female - psi_male,
        V_gap_true = V_female_true - V_male_true
    )

mobility_summary <- df %>%
    group_by(period, female) %>%
    summarise(
        ee_moves = sum(ee),
        ene_moves = sum(ene),
        stayers = sum(ee == 0 & ene == 0),
        ee_rate = mean(ee),
        ene_rate = mean(ene),
        .groups = "drop"
    )

print(mobility_summary)

#--------------------------
# 5.1 gender pay gap (simulated data) and the the impact of Covid
#--------------------------

pay_gap_summary <- wage_long %>%
    group_by(period, female) %>%
    summarise(
        mean_logy = mean(logy),
        sd_logy = sd(logy),
        n = n(),
        .groups = "drop"
    )

pay_gap_did <- lm(logy ~ female * post + age + I(age^2) + I(age^3) + factor(year), data = wage_long)
pay_gap_raw_did <- lm(logy ~ female * post, data = wage_long)

print(pay_gap_summary)
print(summary(pay_gap_did))

#--------------------------
# 5.2 gender gap in firm premium and the the impact of Covid
#--------------------------

akm_premia_wide <- akm_firm_premia_long %>%
    select(firm_id, period, female, psi_hat) %>%
    pivot_wider(
        names_from = female,
        values_from = psi_hat,
        names_prefix = "female_"
    ) %>%
    rename(
        psi_hat_male = female_0,
        psi_hat_female = female_1
    ) %>%
    mutate(
        post = as.integer(period == "post"),
        psi_gap_hat = psi_hat_female - psi_hat_male
    )

firm_premium_gap_summary <- akm_premia_wide %>%
    group_by(period) %>%
    summarise(
        mean_psi_male = mean(psi_hat_male, na.rm = TRUE),
        mean_psi_female = mean(psi_hat_female, na.rm = TRUE),
        mean_psi_gap = mean(psi_gap_hat, na.rm = TRUE),
        sd_psi_gap = sd(psi_gap_hat, na.rm = TRUE),
        n_firms = n(),
        .groups = "drop"
    )

firm_premium_gap_did <- lm(psi_gap_hat ~ post, data = akm_premia_wide)

print(firm_premium_gap_summary)
print(summary(firm_premium_gap_did))

#--------------------------
# 5.3 V_hat and the the impact of Covid
#--------------------------

# Use the iterative Sorkin-style model: recover exp(V), grid-search lambda,
# recover Ve, compute model-consistent EE/EN rates, and iterate.
# For quick debugging only, set rank_method <- "full" to use the stable
# inflow/outflow validation score instead.
rank_method <- "sorkin_full"

rank_male_pre_raw <- estimate_rank(df, period_keep = "pre", gender_keep = 0, method_keep = rank_method)
rank_male_post_raw <- estimate_rank(df, period_keep = "post", gender_keep = 0, method_keep = rank_method)
rank_female_pre_raw <- estimate_rank(df, period_keep = "pre", gender_keep = 1, method_keep = rank_method)
rank_female_post_raw <- estimate_rank(df, period_keep = "post", gender_keep = 1, method_keep = rank_method)

lambda_summary <- bind_rows(
    rank_male_pre_raw,
    rank_male_post_raw,
    rank_female_pre_raw,
    rank_female_post_raw
) %>%
    distinct(period, female, method, lambda_hat) %>%
    mutate(
        gender = ifelse(female == 1, "Female", "Male"),
        # In the current simulation the true outside-offer probability is 0.20
        # in both periods and for both genders. This is useful for checking the
        # recovered lambda, but remember that lambda is estimated from accepted
        # EE moves, so it does not have to match exactly in finite samples.
        lambda_true = 0.20,
        lambda_error = lambda_hat - lambda_true
    )

print(lambda_summary)
write.csv(
    lambda_summary,
    file.path(output_dir, "sorkin_lambda_estimates.csv"),
    row.names = FALSE
)

rank_male_pre <- rank_male_pre_raw %>%
    transmute(firm_id, V_male_pre = V_hat)

rank_male_post <- rank_male_post_raw %>%
    transmute(firm_id, V_male_post = V_hat)

rank_female_pre <- rank_female_pre_raw %>%
    transmute(firm_id, V_female_pre = V_hat)

rank_female_post <- rank_female_post_raw %>%
    transmute(firm_id, V_female_post = V_hat)

rank_all <- rank_male_pre %>%
    full_join(rank_male_post, by = "firm_id") %>%
    full_join(rank_female_pre, by = "firm_id") %>%
    full_join(rank_female_post, by = "firm_id")

require_rows(rank_all, min_rows = 2, label = "rank_all")

rank_coverage <- data.frame(
    measure = c("male_pre", "male_post", "female_pre", "female_post"),
    recovered_firms = c(
        sum(!is.na(rank_all$V_male_pre)),
        sum(!is.na(rank_all$V_male_post)),
        sum(!is.na(rank_all$V_female_pre)),
        sum(!is.na(rank_all$V_female_post))
    )
)

print(rank_coverage)

rank_all <- rank_all %>%
    mutate(
        delta_male = V_male_post - V_male_pre,
        delta_female = V_female_post - V_female_pre,
        gender_gap_pre = V_female_pre - V_male_pre,
        gender_gap_post = V_female_post - V_male_post,
        did_gender_sorting = gender_gap_post - gender_gap_pre
    )

print(summary(rank_all$did_gender_sorting))
print(summary(lm(did_gender_sorting ~ 1, data = rank_all))) # check if the mean of did_gender_sorting is 0

V_hat_gap_summary <- rank_all %>%
    summarise(
        mean_male_pre = mean(V_male_pre, na.rm = TRUE),
        mean_male_post = mean(V_male_post, na.rm = TRUE),
        mean_female_pre = mean(V_female_pre, na.rm = TRUE),
        mean_female_post = mean(V_female_post, na.rm = TRUE),
        mean_gender_gap_pre = mean(gender_gap_pre, na.rm = TRUE),
        mean_gender_gap_post = mean(gender_gap_post, na.rm = TRUE),
        mean_did_gender_sorting = mean(did_gender_sorting, na.rm = TRUE),
        n_firms = n()
    )

print(V_hat_gap_summary)

V_hat_gap_did_data <- rank_all %>%
    select(firm_id, gender_gap_pre, gender_gap_post) %>%
    pivot_longer(
        cols = c(gender_gap_pre, gender_gap_post),
        names_to = "period",
        values_to = "V_gap_hat"
    ) %>%
    mutate(
        period = ifelse(period == "gender_gap_post", "post", "pre"),
        post = as.integer(period == "post")
    )

V_hat_gap_did <- lm(V_gap_hat ~ post, data = V_hat_gap_did_data)

# Sorkin-style sorting component of the gender gap.
# The gender difference in firm premia is evaluated using two price systems:
# male firm premia and female firm premia. This follows the idea that the same
# gender difference in firm allocation can have different wage implications
# depending on whose firm-premium schedule is used.
employment_shares <- wage_long %>%
    count(period, female, firm_id, name = "n") %>%
    group_by(period, female) %>%
    mutate(share = n / sum(n)) %>%
    ungroup() %>%
    select(period, female, firm_id, share) %>%
    pivot_wider(
        names_from = female,
        values_from = share,
        names_prefix = "female_",
        values_fill = 0
    ) %>%
    rename(
        share_male = female_0,
        share_female = female_1
    )

sorkin_gender_gap_by_price <- employment_shares %>%
    left_join(akm_premia_wide, by = c("period", "firm_id")) %>%
    mutate(share_gap = share_female - share_male) %>%
    group_by(period) %>%
    summarise(
        sorting_gap_male_price = sum(share_gap * psi_hat_male, na.rm = TRUE),
        sorting_gap_female_price = sum(share_gap * psi_hat_female, na.rm = TRUE),
        n_firms = n(),
        .groups = "drop"
    ) %>%
    mutate(post = as.integer(period == "post"))

sorkin_male_price_did <- lm(sorting_gap_male_price ~ post, data = sorkin_gender_gap_by_price)
sorkin_female_price_did <- lm(sorting_gap_female_price ~ post, data = sorkin_gender_gap_by_price)

write.csv(
    sorkin_gender_gap_by_price,
    file.path(output_dir, "sorkin_gender_gap_female_male_prices.csv"),
    row.names = FALSE
)

# Sorkin-style opportunity/preference decomposition.
# Preferences are represented by recovered firm values V_hat.
# Opportunities are represented by recovered offer probabilities f_hat.
# For each period, we build four model-implied firm allocations:
#   male opportunities + male preferences,
#   male opportunities + female preferences,
#   female opportunities + male preferences,
#   female opportunities + female preferences.
# Comparing these allocations isolates the role of preferences, opportunities,
# and their interaction in the firm-sorting component of the gender gap.
rank_offer_long <- bind_rows(
    rank_male_pre_raw,
    rank_male_post_raw,
    rank_female_pre_raw,
    rank_female_post_raw
) %>%
    mutate(gender = ifelse(female == 1, "female", "male")) %>%
    select(firm_id, period, gender, V_hat, offer_hat)

make_model_share <- function(offer, value) {
    eps <- 1e-8
    offer_clean <- ifelse(is.finite(offer) & !is.na(offer), offer, eps)
    value_clean <- ifelse(is.finite(value) & !is.na(value), value, 0)

    # A simple Sorkin-style stationary allocation: more offers and higher
    # worker value both increase the share of workers employed at a firm.
    weights <- pmax(offer_clean, eps) * exp(standardize_or_center(value_clean))
    weights / sum(weights, na.rm = TRUE)
}

sorkin_preference_opportunity_data <- rank_offer_long %>%
    pivot_wider(
        names_from = gender,
        values_from = c(V_hat, offer_hat),
        names_sep = "_"
    ) %>%
    left_join(employment_shares, by = c("period", "firm_id")) %>%
    left_join(akm_premia_wide, by = c("period", "firm_id")) %>%
    group_by(period) %>%
    group_modify(~ {
        d <- .x
        d$share_model_mm <- make_model_share(d$offer_hat_male, d$V_hat_male)
        d$share_model_mf <- make_model_share(d$offer_hat_male, d$V_hat_female)
        d$share_model_fm <- make_model_share(d$offer_hat_female, d$V_hat_male)
        d$share_model_ff <- make_model_share(d$offer_hat_female, d$V_hat_female)
        d
    }) %>%
    ungroup()

compute_decomposition_for_price <- function(data, price_var, price_label) {
    data %>%
        group_by(period) %>%
        summarise(
            price_system = price_label,
            observed_sorting_gap = sum((share_female - share_male) * .data[[price_var]], na.rm = TRUE),
            model_sorting_gap = sum((share_model_ff - share_model_mm) * .data[[price_var]], na.rm = TRUE),
            preference_component = sum((share_model_mf - share_model_mm) * .data[[price_var]], na.rm = TRUE),
            opportunity_component = sum((share_model_fm - share_model_mm) * .data[[price_var]], na.rm = TRUE),
            interaction_component = model_sorting_gap - preference_component - opportunity_component,
            n_firms = n(),
            .groups = "drop"
        )
}

sorkin_preference_opportunity <- bind_rows(
    compute_decomposition_for_price(sorkin_preference_opportunity_data, "psi_hat_male", "Male price"),
    compute_decomposition_for_price(sorkin_preference_opportunity_data, "psi_hat_female", "Female price")
) %>%
    arrange(price_system, period)

sorkin_preference_opportunity_change <- sorkin_preference_opportunity %>%
    select(-n_firms) %>%
    pivot_longer(
        cols = c(
            observed_sorting_gap,
            model_sorting_gap,
            preference_component,
            opportunity_component,
            interaction_component
        ),
        names_to = "component",
        values_to = "value"
    ) %>%
    pivot_wider(names_from = period, values_from = value) %>%
    mutate(
        change_post_minus_pre = post - pre
    )

write.csv(
    sorkin_preference_opportunity,
    file.path(output_dir, "sorkin_preference_opportunity_by_period.csv"),
    row.names = FALSE
)

write.csv(
    sorkin_preference_opportunity_change,
    file.path(output_dir, "sorkin_preference_opportunity_change.csv"),
    row.names = FALSE
)

write_sorkin_decomposition_latex_table <- function(decomp_change, path,
                                                   caption = "Sorkin Preference and Opportunity Decomposition",
                                                   digits = 3) {
    component_labels <- c(
        "observed_sorting_gap" = "Observed sorting gap",
        "model_sorting_gap" = "Model sorting gap",
        "preference_component" = "Preferences",
        "opportunity_component" = "Opportunities",
        "interaction_component" = "Interaction"
    )

    fmt <- function(x) ifelse(is.na(x), "", formatC(x, digits = digits, format = "f"))

    table_wide <- decomp_change %>%
        mutate(component = component_labels[component]) %>%
        pivot_wider(
            id_cols = component,
            names_from = price_system,
            values_from = c(pre, post, change_post_minus_pre),
            names_glue = "{price_system}: {.value}"
        )

    desired_cols <- c(
        "component",
        "Male price: pre",
        "Male price: post",
        "Male price: change_post_minus_pre",
        "Female price: pre",
        "Female price: post",
        "Female price: change_post_minus_pre"
    )
    table_wide <- table_wide[, desired_cols]
    names(table_wide) <- c(
        "",
        "Pre",
        "Post",
        "Post-Pre",
        "Pre",
        "Post",
        "Post-Pre"
    )

    for (j in seq_along(table_wide)) {
        if (is.numeric(table_wide[[j]])) {
            table_wide[[j]] <- fmt(table_wide[[j]])
        }
    }

    body <- apply(table_wide, 1, function(row) paste(escape_latex(row), collapse = " & "))
    lines <- c(
        "\\begin{table}[!htbp]\\centering",
        paste0("\\caption{", escape_latex(caption), "}"),
        "\\begin{tabular}{lcccccc}",
        "\\hline",
        " & \\multicolumn{3}{c}{Male price} & \\multicolumn{3}{c}{Female price} \\\\",
        "\\cline{2-4} \\cline{5-7}",
        paste(names(table_wide), collapse = " & "),
        "\\\\ \\hline",
        paste0(body, " \\\\"),
        "\\hline",
        "\\multicolumn{7}{l}{\\footnotesize{Preferences use recovered firm values; opportunities use recovered offer probabilities.}} \\\\",
        "\\multicolumn{7}{l}{\\footnotesize{Post-Pre is the change in the decomposition after Covid relative to before Covid.}} \\\\",
        "\\end{tabular}",
        "\\end{table}"
    )
    writeLines(lines, path)
}

write_sorkin_decomposition_latex_table(
    sorkin_preference_opportunity_change,
    file.path(output_dir, "sorkin_preference_opportunity_table.tex")
)

#--------------------------
# 5.4 V_true (simulated data) and the the impact of Covid
#--------------------------

V_true_gap_summary <- firm_gap %>%
    group_by(period) %>%
    summarise(
        mean_V_male_true = mean(V_male_true),
        mean_V_female_true = mean(V_female_true),
        mean_V_gap_true = mean(V_gap_true),
        sd_V_gap_true = sd(V_gap_true),
        n_firms = n(),
        .groups = "drop"
    )

V_true_gap_did <- lm(V_gap_true ~ post, data = firm_gap)

print(V_true_gap_summary)
print(summary(V_true_gap_did))

extract_did_row <- function(model, term, outcome_label) {
    coefs <- summary(model)$coefficients
    if (!term %in% rownames(coefs)) {
        return(data.frame(
            outcome = outcome_label,
            did_term = term,
            estimate = NA_real_,
            std_error = NA_real_,
            p_value = NA_real_,
            n = nobs(model)
        ))
    }
    data.frame(
        outcome = outcome_label,
        did_term = term,
        estimate = coefs[term, "Estimate"],
        std_error = coefs[term, "Std. Error"],
        p_value = coefs[term, "Pr(>|t|)"],
        n = nobs(model)
    )
}

# Adds the conventional significance stars used in regression tables.
significance_stars <- function(p_value) {
    if (is.na(p_value)) {
        return("")
    }
    if (p_value < 0.001) {
        return("***")
    }
    if (p_value < 0.01) {
        return("**")
    }
    if (p_value < 0.05) {
        return("*")
    }
    if (p_value < 0.1) {
        return("+")
    }
    ""
}

# Writes a regression-style LaTeX table like the one used in the report:
# coefficients are stacked over standard errors, model fit statistics are at
# the bottom, and the DID term is displayed alongside the controls.
write_did_regression_latex_table <- function(models, path, caption = "Regression Results",
                                             term_labels = NULL, year_fe = NULL,
                                             digits = 3) {
    model_names <- paste0("(", seq_along(models), ")")
    all_terms <- unique(unlist(lapply(models, function(model) rownames(summary(model)$coefficients))))
    all_terms <- all_terms[all_terms != "(Intercept)"]

    preferred_order <- c(
        "female",
        "post",
        "female:post",
        "age",
        "I(age^2)",
        "I(age^3)"
    )
    term_order <- c(preferred_order[preferred_order %in% all_terms], setdiff(all_terms, preferred_order))

    if (is.null(term_labels)) {
        term_labels <- term_order
        names(term_labels) <- term_order
    }
    if (is.null(year_fe)) {
        year_fe <- rep(FALSE, length(models))
    }

    fmt_num <- function(x) {
        ifelse(is.na(x), "", formatC(x, digits = digits, format = "f"))
    }

    table_rows <- list()
    for (term in term_order) {
        coef_row <- c(ifelse(term %in% names(term_labels), term_labels[[term]], term))
        se_row <- c("")

        for (model in models) {
            coefs <- summary(model)$coefficients
            if (term %in% rownames(coefs)) {
                estimate <- coefs[term, "Estimate"]
                std_error <- coefs[term, "Std. Error"]
                p_value <- coefs[term, "Pr(>|t|)"]
                coef_row <- c(coef_row, paste0(fmt_num(estimate), significance_stars(p_value)))
                se_row <- c(se_row, paste0("(", fmt_num(std_error), ")"))
            } else {
                coef_row <- c(coef_row, "")
                se_row <- c(se_row, "")
            }
        }

        table_rows[[length(table_rows) + 1]] <- coef_row
        table_rows[[length(table_rows) + 1]] <- se_row
    }

    add_stat_row <- function(label, values) {
        c(label, values)
    }

    nobs_row <- add_stat_row("Num.Obs.", formatC(sapply(models, nobs), format = "d"))
    r2_row <- add_stat_row("R2", fmt_num(sapply(models, function(model) summary(model)$r.squared)))
    adj_r2_row <- add_stat_row("R2 Adj.", fmt_num(sapply(models, function(model) summary(model)$adj.r.squared)))
    aic_row <- add_stat_row("AIC", fmt_num(sapply(models, AIC)))
    bic_row <- add_stat_row("BIC", fmt_num(sapply(models, BIC)))
    rmse_row <- add_stat_row("RMSE", fmt_num(sapply(models, function(model) sqrt(mean(resid(model)^2, na.rm = TRUE)))))
    year_fe_row <- add_stat_row("FE: year", ifelse(year_fe, "x", ""))

    table_rows <- c(
        table_rows,
        list(nobs_row, r2_row, adj_r2_row, aic_row, bic_row, rmse_row, year_fe_row)
    )

    body <- vapply(
        table_rows,
        function(row) paste(escape_latex(row), collapse = " & "),
        character(1)
    )

    lines <- c(
        "\\begin{table}[!htbp]\\centering",
        paste0("\\caption{", escape_latex(caption), "}"),
        paste0("\\begin{tabular}{l", paste(rep("c", length(models)), collapse = ""), "}"),
        "\\hline",
        paste(c("", model_names), collapse = " & "),
        "\\\\ \\hline",
        paste0(body, " \\\\"),
        "\\hline",
        "\\multicolumn{", length(models) + 1, "}{l}{\\footnotesize{+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}} \\\\",
        "\\end{tabular}",
        "\\end{table}"
    )
    writeLines(lines, path)
}

did_results_table <- bind_rows(
    extract_did_row(pay_gap_raw_did, "female:post", "Raw log wage gender gap"),
    extract_did_row(pay_gap_did, "female:post", "Adjusted log wage gender gap"),
    extract_did_row(firm_premium_gap_did, "post", "AKM firm-premium gender gap"),
    extract_did_row(V_hat_gap_did, "post", "Recovered firm-value gender gap"),
    extract_did_row(V_true_gap_did, "post", "True firm-value gender gap"),
    extract_did_row(sorkin_male_price_did, "post", "Sorkin sorting gap, male price"),
    extract_did_row(sorkin_female_price_did, "post", "Sorkin sorting gap, female price")
)

print(did_results_table)

write.csv(
    did_results_table,
    file.path(output_dir, "did_results_table.csv"),
    row.names = FALSE
)

write_simple_latex_table(
    did_results_table,
    file.path(output_dir, "did_results_summary_table.tex"),
    caption = "Difference-in-Differences Summary"
)

rank_worker_long <- rank_offer_long %>%
    mutate(
        gender = ifelse(gender == "female", "Female", "Male"),
        female = as.integer(gender == "Female")
    ) %>%
    select(firm_id, period, female, V_hat)

did_worker_regression_data <- wage_long %>%
    left_join(
        akm_firm_premia_long %>% select(period, female, firm_id, psi_hat),
        by = c("period", "female", "firm_id")
    ) %>%
    left_join(
        rank_worker_long,
        by = c("period", "female", "firm_id")
    )

firm_premium_worker_did <- lm(psi_hat ~ female * post, data = did_worker_regression_data)
firm_premium_worker_year_did <- lm(psi_hat ~ female * post + factor(year), data = did_worker_regression_data)

did_regression_models <- list(
    "log(hourly wage)" = pay_gap_raw_did,
    "log(hourly wage)" = pay_gap_did,
    "firm premia" = firm_premium_worker_did,
    "firm premia" = firm_premium_worker_year_did
)

did_regression_extra_rows <- data.frame(
    term = c("Std.Errors", "FE: year"),
    `log(hourly wage)` = c("by: worker_id", ""),
    `log(hourly wage)` = c("by: worker_id", "x"),
    `firm premia` = c("by: worker_id", ""),
    `firm premia` = c("by: worker_id", "x"),
    check.names = FALSE
)

modelsummary::modelsummary(
    did_regression_models,
    output = file.path(output_dir, "did_results_table.tex"),
    title = "Regression Results",
    coef_map = c(
        "female" = "female",
        "post" = "post",
        "female:post" = "female-post",
        "age" = "age",
        "I(age^2)" = "age2",
        "I(age^3)" = "age3"
    ),
    vcov = list(
        ~worker_id,
        ~worker_id,
        ~worker_id,
        ~worker_id
    ),
    stars = c("+" = 0.1, "*" = 0.05, "**" = 0.01, "***" = 0.001),
    gof_map = c("nobs", "r.squared", "adj.r.squared", "aic", "bic", "rmse"),
    add_rows = did_regression_extra_rows,
    notes = "+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"
)

#--------------------------
# 5.5 check the correlation of V_hat between male and female, pre and post sepearately
#--------------------------

V_hat_correlations <- data.frame(
    period = c("pre", "post"),
    corr_male_female = c(
        cor(rank_all$V_male_pre, rank_all$V_female_pre, use = "complete.obs"),
        cor(rank_all$V_male_post, rank_all$V_female_post, use = "complete.obs")
    ),
    n_complete = c(
        sum(complete.cases(rank_all$V_male_pre, rank_all$V_female_pre)),
        sum(complete.cases(rank_all$V_male_post, rank_all$V_female_post))
    )
)

print(V_hat_correlations)

#--------------------------
# 5.6 check the correlation of V_true between male and female, pre and post sepearately
#--------------------------

V_true_correlations <- firm_gap %>%
    group_by(period) %>%
    summarise(
        corr_male_female = cor(V_male_true, V_female_true),
        n_firms = n(),
        .groups = "drop"
    )

print(V_true_correlations)
```

# 6. Visualization (for both true simulated data and estimated data)
```{r visualization, echo=TRUE}
#--------------------------
# 6.0 data preparation
#--------------------------

# rank_all to long format
rank_all_long <- rank_all %>%
    select(
        firm_id,
        V_male_pre,
        V_male_post,
        V_female_pre,
        V_female_post
    ) %>%
    tidyr::pivot_longer(
        cols = -firm_id,
        names_to = c("gender", "period"),
        names_pattern = "V_(male|female)_(pre|post)",
        values_to = "V_hat"
    )

# df_simulated to long format
df_simulated_male <- df_simulated %>%
    select(
        firm_id, period,
        V_male_true,
        psi_male
    ) %>%
    mutate(gender = "male") %>%
    rename(
        V_true = V_male_true,
        psi = psi_male
    )

df_simulated_female <- df_simulated %>%
    select(
        firm_id, period,
        V_female_true,
        psi_female
    ) %>%
    mutate(gender = "female") %>%
    rename(
        V_true = V_female_true,
        psi = psi_female
    )

df_simulated_long <- bind_rows(df_simulated_male, df_simulated_female)

# merge estimated ranks with true simulated firm values
comparison <- rank_all_long %>%
    left_join(
        df_simulated_long,
        by = c("firm_id", "period", "gender")
    ) %>%
    left_join(
        akm_firm_premia_long %>%
            mutate(gender = ifelse(female == 1, "female", "male")) %>%
            select(firm_id, period, gender, psi_hat),
        by = c("firm_id", "period", "gender")
    )

require_rows(
    comparison %>% filter(!is.na(V_hat), !is.na(V_true)),
    min_rows = 2,
    label = "comparison data for validity plots"
)

# sanity check:
# 1.how many firms in each gender-period?
print(table(comparison$gender, comparison$period))
# 2. any missing value especially in V_hat?
print(colSums(is.na(comparison)))

psi_validation_plot_data <- comparison %>%
    filter(!is.na(psi), !is.na(psi_hat)) %>%
    group_by(gender, period) %>%
    mutate(
        # AKM firm effects are normalized within gender-period cells and can
        # contain finite-sample outliers when identification comes from a small
        # mover set. Rank-normalizing both axes makes the plot a cleaner check
        # of whether AKM recovers the ordering of true firm premia.
        psi_true_rank = rank_to_normal_score(psi),
        psi_hat_rank = rank_to_normal_score(psi_hat)
    ) %>%
    ungroup() %>%
    transmute(
        firm_id,
        gender,
        period,
        psi_true_rank,
        psi_hat_rank
    )

psi_validation_summary <- psi_validation_plot_data %>%
    group_by(gender, period) %>%
    summarise(
        corr_rank_true_recovered_psi = cor(psi_hat_rank, psi_true_rank, use = "complete.obs"),
        n_firms = n(),
        .groups = "drop"
    )

print(psi_validation_summary)

write.csv(
    psi_validation_summary,
    file.path(output_dir, "validity_check_firm_premia_summary.csv"),
    row.names = FALSE
)

write.csv(
    psi_validation_plot_data,
    file.path(output_dir, "validity_check_firm_premia_plot_data.csv"),
    row.names = FALSE
)

# Estimated firm premia to long format for the density plot below.
firm_premia_long <- akm_firm_premia_long %>%
    transmute(
        firm_id,
        period,
        gender,
        psi = psi_hat
    )

# V_true and V_hat distribution data
value_dist_long <- comparison %>%
    filter(!is.na(V_hat), !is.na(V_true)) %>%
    transmute(
        firm_id,
        gender = ifelse(gender == "female", "Female", "Male"),
        period,
        `Recovered value: V_hat` = V_hat,
        `True value: V_true` = V_true
    ) %>%
    pivot_longer(
        cols = c(`Recovered value: V_hat`, `True value: V_true`),
        names_to = "value_type",
        values_to = "value"
    )
#--------------------------
# 6.1 Check for validity of our measure based on our simulation
#--------------------------

# total sample
# Report both the raw correlation and the rank-normalized plotting correlation.
# The graph should never use raw PageRank/Sorkin magnitudes directly because a
# few weakly connected firms can stretch the axis and destroy the visual check.
validity_raw <- cor(comparison$V_hat, comparison$V_true, use = "complete.obs")

print(paste0(
    "the raw correlation between full-model V_hat and V_true for total sample: ",
    validity_raw
))

comparison_plot <- comparison %>%
    filter(!is.na(V_hat), !is.na(V_true)) %>%
    transmute(
        firm_id,
        # Rank-normalizing both axes prevents a few finite-sample PageRank
        # outliers from stretching the figure while preserving the ordering.
        hat = rank_to_normal_score(V_hat),
        true = rank_to_normal_score(V_true)
    )

require_rows(comparison_plot, min_rows = 2, label = "total-sample validity plot data")

validity_rank <- cor(comparison_plot$hat, comparison_plot$true, use = "complete.obs")
print(paste0(
    "the rank-normalized plotting correlation between V_hat and V_true for total sample: ",
    validity_rank
))

write.csv(
    comparison_plot,
    file.path(output_dir, "validity_check_plot_data.csv"),
    row.names = FALSE
)

line_range <- range(c(comparison_plot$hat, comparison_plot$true), finite = TRUE)
line_45 <- data.frame(
    x = line_range[1],
    xend = line_range[2],
    y = line_range[1],
    yend = line_range[2],
    line_type = "45-degree line"
)

p1 <- ggplot(comparison_plot, aes(x = hat, y = true)) +
    geom_segment(
        data = line_45,
        aes(x = x, y = y, xend = xend, yend = yend, linetype = line_type),
        color = "grey40",
        linewidth = 0.8,
        inherit.aes = FALSE,
        show.legend = TRUE
    ) +
    geom_point(color = "grey60", size = 0.75, alpha = 0.50, show.legend = FALSE) +
    geom_smooth(
        aes(linetype = "Linear fit"),
        method = "lm",
        se = FALSE,
        color = "black",
        linewidth = 1.1,
        show.legend = TRUE
    ) +
    scale_linetype_manual(
        name = "Line",
        breaks = c("45-degree line", "Linear fit"),
        labels = c("45 degree", "Fitted line"),
        values = c(
            "45-degree line" = "dashed",
            "Linear fit" = "solid"
        )
    ) +
    labs(
        x = "Recovered firm value (standardized V_hat)",
        y = "True latent firm value (standardized V_true)",
        title = "Validity check of revealed-preference ranking (total sample)"
    ) +
    coord_equal() +
    theme_classic(base_size = 14) +
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom"
    )

validity_check_path <- file.path(output_dir, "validity_check.png")
ggsave(validity_check_path, plot = p1, width = 8, height = 6, dpi = 300)
print(paste0("Saved: ", normalizePath(validity_check_path, mustWork = FALSE)))

# gender_period subsamples
validity_sub <- comparison %>%
    group_by(gender, period) %>%
    summarise(
        raw_corr = cor(V_hat, V_true, use = "complete.obs"),
        n = sum(complete.cases(V_hat, V_true)),
        .groups = "drop"
    )
print("the correlation between V_hat and V_true for gender_period subsamples: ")
print(validity_sub)

comparison_plot_sub <- comparison %>%
    filter(!is.na(V_hat), !is.na(V_true)) %>%
    group_by(gender, period) %>%
    mutate(
        # Rank-normalize within each panel, since the plot is faceted by
        # gender-period and each recovered ranking is estimated separately.
        hat = rank_to_normal_score(V_hat),
        true = rank_to_normal_score(V_true)
    ) %>%
    ungroup() %>%
    select(firm_id, gender, period, hat, true)

require_rows(comparison_plot_sub, min_rows = 2, label = "subsample validity plot data")

validity_sub_rank <- comparison_plot_sub %>%
    group_by(gender, period) %>%
    summarise(
        rank_plot_corr = cor(hat, true, use = "complete.obs"),
        n = sum(complete.cases(hat, true)),
        .groups = "drop"
    )
print("the rank-normalized plotting correlation between V_hat and V_true for gender_period subsamples: ")
print(validity_sub_rank)

write.csv(
    comparison_plot_sub,
    file.path(output_dir, "validity_check_sub_plot_data.csv"),
    row.names = FALSE
)

line_45_sub <- comparison_plot_sub %>%
    group_by(gender, period) %>%
    summarise(
        x = min(c(hat, true), na.rm = TRUE),
        xend = max(c(hat, true), na.rm = TRUE),
        y = x,
        yend = xend,
        line_type = "45-degree line",
        .groups = "drop"
    )

p1_sub <- ggplot(comparison_plot_sub, aes(x = hat, y = true)) +
    geom_segment(
        data = line_45_sub,
        aes(x = x, y = y, xend = xend, yend = yend, linetype = line_type),
        color = "grey40",
        linewidth = 0.8,
        inherit.aes = FALSE,
        show.legend = TRUE
    ) +
    geom_point(color = "grey60", size = 0.75, alpha = 0.50, show.legend = FALSE) +
    geom_smooth(
        aes(linetype = "Linear fit"),
        method = "lm",
        se = FALSE,
        color = "black",
        linewidth = 1.1,
        show.legend = TRUE
    ) +
    facet_grid(gender ~ period) +
    scale_linetype_manual(
        name = "Line",
        breaks = c("45-degree line", "Linear fit"),
        labels = c("45 degree", "Fitted line"),
        values = c(
            "45-degree line" = "dashed",
            "Linear fit" = "solid"
        )
    ) +
    labs(
        x = "Recovered firm value (standardized V_hat)",
        y = "True latent firm value (standardized V_true)",
        title = "Validity check of revealed-preference ranking (gender_period subsamples)"
    ) +
    coord_equal() +
    theme_classic(base_size = 14) +
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom"
    )

validity_check_sub_path <- file.path(output_dir, "validity_check_sub.png")
ggsave(validity_check_sub_path, plot = p1_sub, width = 8, height = 6, dpi = 300)
print(paste0("Saved: ", normalizePath(validity_check_sub_path, mustWork = FALSE)))

#--------------------------
# 6.1.1 Validity check for AKM-recovered firm premia
#--------------------------

psi_line_45 <- psi_validation_plot_data %>%
    group_by(gender, period) %>%
    summarise(
        x = min(c(psi_hat_rank, psi_true_rank), na.rm = TRUE),
        xend = max(c(psi_hat_rank, psi_true_rank), na.rm = TRUE),
        y = x,
        yend = xend,
        line_type = "45-degree line",
        .groups = "drop"
    )

p_psi_validity <- ggplot(psi_validation_plot_data, aes(x = psi_hat_rank, y = psi_true_rank)) +
    geom_segment(
        data = psi_line_45,
        aes(x = x, y = y, xend = xend, yend = yend, linetype = line_type),
        color = "grey40",
        linewidth = 0.8,
        inherit.aes = FALSE,
        show.legend = TRUE
    ) +
    geom_point(color = "grey60", size = 0.75, alpha = 0.50, show.legend = FALSE) +
    geom_smooth(
        aes(linetype = "Linear fit"),
        method = "lm",
        se = FALSE,
        color = "black",
        linewidth = 1.1,
        show.legend = TRUE
    ) +
    facet_grid(gender ~ period) +
    scale_linetype_manual(
        name = "Line",
        breaks = c("45-degree line", "Linear fit"),
        labels = c("45 degree", "Fitted line"),
        values = c(
            "45-degree line" = "dashed",
            "Linear fit" = "solid"
        )
    ) +
    labs(
        x = "Recovered firm premium (rank-normalized AKM psi_hat)",
        y = "True simulated firm premium (rank-normalized psi)",
        title = "Validity check of AKM-recovered firm premia"
    ) +
    coord_equal() +
    theme_classic(base_size = 14) +
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom"
    )

ggsave(
    filename = file.path(output_dir, "validity_check_firm_premia.png"),
    plot = p_psi_validity,
    width = 8,
    height = 6,
    dpi = 300
)

#--------------------------
# 6.1.2 Diagnostic check for recovered lambda
#--------------------------

p_lambda <- ggplot(lambda_summary, aes(x = period, y = lambda_hat, color = gender, group = gender)) +
    geom_hline(
        aes(yintercept = lambda_true, linetype = "True lambda"),
        color = "grey40",
        linewidth = 0.8,
        show.legend = TRUE
    ) +
    geom_point(size = 2.2) +
    geom_line(linewidth = 0.8) +
    scale_linetype_manual(
        name = "Line",
        values = c("True lambda" = "dashed")
    ) +
    labs(
        title = "Diagnostic check of recovered lambda",
        x = "Period",
        y = "Outside-offer probability lambda",
        color = "Gender"
    ) +
    theme_classic(base_size = 14) +
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom"
    )

ggsave(
    filename = file.path(output_dir, "validity_check_lambda.png"),
    plot = p_lambda,
    width = 7,
    height = 5,
    dpi = 300
)

#--------------------------
# 6.2 wage distribution by gender and period
#--------------------------

p_wage_dist <- ggplot(wage_long, aes(x = logy, color = gender, fill = gender)) +
    geom_density(alpha = 0.25, linewidth = 0.8) +
    facet_wrap(~period) +
    labs(
        title = "Distribution of Log Wages by Gender and Period",
        x = "Log wage",
        y = "Density",
        color = "Gender",
        fill = "Gender"
    ) +
    theme_minimal()

ggsave(
    filename = file.path(output_dir, "distribution_log_wage_gender_period.png"),
    plot = p_wage_dist,
    width = 8,
    height = 5,
    dpi = 300
)

# ------------------------------------------------------------
# 6.3. Firm wage-premia distribution by gender and period
# ------------------------------------------------------------

p_premia_dist <- ggplot(firm_premia_long, aes(x = psi, color = gender, fill = gender)) +
    geom_density(alpha = 0.25, linewidth = 0.8) +
    facet_wrap(~period) +
    labs(
        title = "Distribution of Firm Wage Premia by Gender and Period",
        x = "Firm wage premium (psi)",
        y = "Density",
        color = "Gender",
        fill = "Gender"
    ) +
    theme_minimal()

ggsave(
    filename = file.path(output_dir, "distribution_firm_premia_gender_period.png"),
    plot = p_premia_dist,
    width = 8,
    height = 6,
    dpi = 300
)

# ------------------------------------------------------------
# 6.4. V_true and V_hat distribution by gender and period
# ------------------------------------------------------------

p_value_dist <- ggplot(value_dist_long, aes(x = value, color = gender, fill = gender)) +
    geom_density(alpha = 0.25, linewidth = 0.8) +
    facet_grid(value_type ~ period, scales = "free") +
    labs(
        title = "Distribution of True and Recovered Firm Values by Period",
        x = "Firm value",
        y = "Density",
        color = "Gender",
        fill = "Gender"
    ) +
    theme_minimal()

ggsave(
    filename = file.path(output_dir, "distribution_values_true_hat_gender_period.png"),
    plot = p_value_dist,
    width = 9,
    height = 6,
    dpi = 300
)

premia_summary_table <- firm_premia_long %>%
    group_by(gender, period) %>%
    summarise(
        n_firms = n(),
        mean = mean(psi, na.rm = TRUE),
        sd = sd(psi, na.rm = TRUE),
        p10 = quantile(psi, 0.10, na.rm = TRUE),
        p50 = quantile(psi, 0.50, na.rm = TRUE),
        p90 = quantile(psi, 0.90, na.rm = TRUE),
        .groups = "drop"
    )

print(premia_summary_table)


modelsummary::datasummary_df(
    premia_summary_table,
    output = file.path(output_dir, "premia_summary_statistics.tex"),
    title = "Firm Premia Summary Statistics by Gender and Period"
)

# Check exported files
list.files(output_dir)

#--------------------------
# 6.5 Test for compensating differentials
#--------------------------

comparison_cd <- comparison %>%
    transmute(
        firm_id,
        gender,
        period,
        hat = V_hat,
        psi = psi_hat
    ) %>%
    filter(!is.na(hat), !is.na(psi))

require_rows(comparison_cd, min_rows = 2, label = "compensating-differentials plot data")

print(summary(lm(psi ~ hat, data = comparison_cd)))

p2 <- ggplot(comparison_cd, aes(x = hat, y = psi)) +
    geom_point(color = "grey60", size = 1.1, alpha = 0.55) +
    geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 1.1) +
    facet_grid(gender ~ period) +
    labs(
        x = "Recovered firm value (V_hat)",
        y = "Firm wage premium (psi)",
        title = "Compensating differentials"
    ) +
    theme_classic(base_size = 14) +
    theme(plot.title = element_text(hjust = 0.5))

compensating_differentials_path <- file.path(output_dir, "compensating_differentials.png")
ggsave(compensating_differentials_path, plot = p2, width = 8, height = 6, dpi = 300)
print(paste0("Saved: ", normalizePath(compensating_differentials_path, mustWork = FALSE)))

```
